---
title: "DevlopingDataProducts"
author: "Anita Gaikwad"
date: "9 September 2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

    This presentation was created as part of the Developing Data Products course of the Coursera Data Science Specialization.

    The goal of the project is to create a web page presentation using R Markdown that features a plot created with Plotly, and to host the resulting web page on either GitHub Pages, RPubs, or NeoCities.

    The interactive plot on the next slide represents the number of road accidents in Great Britain from 2005 to 2015, grouped by severity (slight, serious, or fatal).

##Obtaining Data

Two data sets are used for this presentation

    Accidents_2015.csv from Road Safety - Accidents 2015

    Accidents0514.csv from All STATS19 data (accident, casualties and vehicle tables) for 2005 to 2014

These two data sets, alongwith many other similar datasets, can be found in the Road Safety Data data directory provided by the Department of Transport of the U.K. government.


##Loading required packages and data
```{r}
rm(list=ls())
library(plotly)
library(data.table)
library(tidyr)
library(lubridate)
library(zoo)
library(dplyr)
#packages <- c("plotly", "data.table", "tidyr",
#              "lubridate", "zoo", "dplyr")
#lapply(packages, require, character.only = TRUE)

accidents0514 <- fread("F:/Anita RJ/DataScience/Coursera/DevlopingDataProducts/Week3Assign/Accidents0514.csv",
                       header = TRUE, sep = ",")
accidents0514 <- accidents0514 %>% select(Accident_Severity, Date)

accidents <- fread("F:/Anita RJ/DataScience/Coursera/DevlopingDataProducts/Week3Assign/Accidents_2015.csv",
                     header = TRUE, sep = ",")
accidents <- accidents %>% select(Accident_Severity, Date)
```

##Data preparation
```{r}
# convert severity to factor and add labels
accidents$Accident_Severity <- 
        factor(accidents$Accident_Severity, levels = 1:3,
               labels = c("Fatal", "Serious", "Slight"))

# convert date strings to Date objects
accidents$Date <- dmy(accidents$Date)

# group data by date and severity, get count, one row per date
accident_count <- accidents %>% group_by(Date, Accident_Severity) %>%
        summarise(count = n()) %>%
        spread(key = Accident_Severity, value = count) %>% 
        as.data.frame()

# create a smoother for each severity to visualise general trends
loess_slight <- loess(Slight~as.numeric(Date), data = accident_count)
loess_serious <- loess(Serious~as.numeric(Date),data = accident_count)
loess_fatal <- loess(Fatal~as.numeric(Date), data = accident_count)
```

##Interactive plotting using plotly
```{r}
# plot data
plot_ly(accident_count) %>%
    add_trace(x = ~Date, y = ~Slight, type="scatter",
              mode = "markers", name = "slight",
              legendgroup = "slight",
              marker = list(color = "#52A9BD")) %>%
    add_trace(x = ~Date, y = ~Serious, type="scatter",
              mode = "markers", name = "serious",
              legendgroup = "serious", 
              marker = list(color = "#FFF16B")) %>%
    add_trace(x = ~Date, y = ~Fatal, type="scatter",
              mode = "markers", name = "fatal",
              legendgroup = "fatal", 
              marker = list(color = "#F5677D")) %>%
  add_trace(x = as.Date(loess_slight$x), y = fitted(loess_slight),
              type="scatter", mode = "lines",
              line = list(color = '#1A7A90'), 
              name = "slight Loess smoother",legendgroup = "slight", 
              hoverinfo = 'none', showlegend = FALSE) %>%
    add_trace(x = as.Date(loess_serious$x), y = fitted(loess_serious),
              type = "scatter", mode = "lines",
              line = list(color = '#E9D625'),
              name = "serious Loess smoother",legendgroup = "serious",
              hoverinfo = 'none', showlegend = FALSE) %>%
    add_lines(x = as.Date(loess_fatal$x), y = fitted(loess_fatal),
              type="scatter", mode = "lines",
              line = list(color = '#DC2340'),
              name = "fatal Loess smoother", legendgroup = "fatal",
              hoverinfo = 'none', showlegend = FALSE) %>%
    layout(xaxis = list(title = "Year"),
           yaxis = list(title = "Number of Accidents")
    )
```

